An Architecture for Real-Time Location Recommendation for Field Data Collection

### 2017-03-09 Dynamic Enumerator Allocation/Paper/v2/Paper.tex
Keep paper updated

\usepackage{hyperref}

\usepackage{minted}
\usepackage{tcolorbox}
\usepackage{etoolbox}
\BeforeBeginEnvironment{minted}{\begin{tcolorbox}}%
\AfterEndEnvironment{minted}{\end{tcolorbox}}%
\AtBeginEnvironment{minted}{\fontsize{8}{8}}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Real-Time Location Recommendation for Field Data Collection}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Field data collection is one of the main activities performed by the statistical agencies of a country. Data collection activities have a common workflow with Multi Depot Vehicle Routing Problem (MDVRP). The use MDVRP to generate pre-calculated routes resulted in a total route costs with high standard deviation. Real-time mechanism by utilizing the paradigm of publish / subscribe combined with COEs MDVRP algorithm is proposed to reduce the inequality (large variation) the time of completion. The test results show that the mechanisms pub / sub combined with COEs produce a total route times more prevalent among enumerators compared with pre-calculated COEs algorithm.
\end{abstract}

@@

%-----------------------------------------------------------------------------%
\section{Introduction}
\label{sec:introduction}
%-----------------------------------------------------------------------------%
Pengumpulan data lapangan merupakan salah satu tugas dan wewenang dari lembaga statistik suatu negara. Terdapat dua macam metode pengumpulan data yang digunakan, yaitu: pengumpulan data primer dan pengumpulan data sekunder. Pada pengumpulan data primer, petugas pengumpulan data secara langsung melakukan interview dengan responden, sementara pada pengumpulan data sekunder, instansi melakukan kompilasi data yang telah dilakukan oleh pihak lain. Pada proses pengumpulan data (primer), petugas pengumpulan data dialokasikan pada sejumlah lokasi pencacahan yang telah ditentukan. Mekanisme pengalokasian yang biasanya digunakan adalah alokasi dengan jumlah lokasi yang sama antar petugas.

Alur kerja pada proses pengumpulan adalah, setiap petugas pengumpulan data memulai pengumpulan data dari sebuah titik tertentu, baik itu kantor maupun tempat tinggal masing-masing petugas. Kemudian petugas melakukan pengumpulan data pada sebuah lokasi sampai selesai, kemudian dilanjutkan dengan lokasi berikutnya, dan seterusnya sampai seluruh lokasi pencacahan dikunjungi. Alur kerja pengumpulan data ini sangat mirip dengan \textit{Vehicle Routing Problem} (VRP), atau lebih spesifik lagi \textit{Multi Depot} VRP.

Akan tetapi, meskipun masalah pengumpulan data memiliki kesamaan alur kerja dengan MDVRP, akan tetapi implementasi algoritma MDVRP secara mentah tidak dapat menjadi solusi atas permasalahan alokasi petugas pengumpulan data, karena salah satu komponen yang penting, yaitu lamanya waktu pelayanan, tidak tersedia sampai dengan lokasi dikunjungi. This research aims to propose an integration antara MDVRP dengan mekanisme pub/sub untuk menciptakan sistem rekomendasi lokasi pencacahan secara real-time.

%-----------------------------------------------------------------------------%
\section{Literature Review}
\label{sec:literature-review}
%-----------------------------------------------------------------------------%
%-----------------------------------------------------------------------------%
\subsection{Multi-Depot Vehicle Routing Problem}
\label{ssec:mdvrp}
%-----------------------------------------------------------------------------%
\textit{Multi-Depot Vehicle Routing Problem (MDVRP)} merupakan salah satu varian dari VRP klasik dimana terdapat lebih dari satu \textit{depot} yang digunakan \cite{montoya-torres_literature_2015}. Gambar \ref{fig:mdvrp-illustration} menunjukkan contoh solusi dari permasalahan VRP yang menggunakan dua depot dan dua buah rute kendaraan yang berasosiasi dengan masing-masing depot. Basically, a solution to this problem is a set of vehicle routes such that: (i) each vehicle route starts and ends at the same depot, (ii) each customer is served exactly once by one vehicle, (iii) the total demand on each route does not exceed vehicle capacity (iv) the maximum route time is satisfied and (v) the total cost is minimized.

\begin{figure}[h]
	\centering
	\includegraphics[width=5cm]{Resources/Images/mdvrp-illustration}
	\caption{Ilustrasi \textit{Multi Depot} VRP}
	\label{fig:mdvrp-illustration}
\end{figure}

The MDVRP can be formalized as follows. Let $G = (V, A)$ be a complete graph, where $V$ is the set of nodes and $A$ is the set of arcs. The nodes are partitioned into two subsets: the customers to be served, $V_C$ = \{1,..., $N$\}, and the multiple depots $V_D$ = \{$N$+1,..., $N+M$\}, with $V_C \cup V_D$ = $V$ and $V_C \cap V_D$ = $\oslash$. There is a non-negative cost $c_ij$ associated with each arc $(i, j) \in A$. The demand of each customer is $d_i$ (there is no demand at the depot nodes). There is also a fleet of $K$ identical vehicles, each with capacity $Q$. The service time at each customer $i$ is $t_i$ while the maximum route duration time is set to $T$. $A$ conversion factor $w_{ij}$ might be needed to transform the cost $c_{ij}$ into time units. In this work, however, the cost is the same as the time and distance units, so $w_{ij}$ = $1$.

In the mathematical formulation that follows, binary variables $x_{ijk}$ are equal to $1$ when vehicle $k$ visits node $j$ immediately after node $i$. Auxiliary variables $y_i$ are also used in the subtour elimination constraints.

\begin{equation}
\label{eq:1}
\begin{aligned}
\sum_{i=1}^{N+M}\sum_{j=1}^{N+M}\sum_{k=1}^{K}c_{ij}x_{ijk};
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:2}
\begin{aligned}
\sum_{i=1}^{N+M}\sum_{k=1}^{K}x_{ijk} = 1  (j=1,..., N);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:3}
\begin{aligned}
\sum_{j=1}^{N+M}\sum_{k=1}^{K}x_{ijk} = 1  (j=1,..., N);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:4}
\begin{aligned}
\sum_{i=1}^{N+M}x_{ihk} - \sum_{j=1}^{N+M}x_{hjk} = 0 \\
(k=1,...,K; h=1,...,N+M);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:5}
\begin{aligned}
\sum_{i=1}^{N+M} \sum_{j=1}^{N+M} d_ix_{ijk} \leq Q (k=1,...,K);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:6}
\begin{aligned}
\sum_{i=1}^{N+M} \sum_{j=1}^{N+M} (c_{ij}w_{ij} + t_i) x_{ijk} \leq T (k=1,...,K);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:7}
\begin{aligned}
\sum_{i=N+1}^{N+M} \sum_{j=1}^{N} x_{ijk} \leq 1 (k=1,...,K);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:8}
\begin{aligned}
\sum_{j=N+1}^{N+M} \sum_{i=1}^{N} x_{ijk} \leq 1 (k=1,...,K);
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:9}
\begin{aligned}
y_i - y_j + (M + N)x_{ijk} \leq N + M - 1; \\
for 1 \leq i \neq j \leq N and 1 \leq k \leq K;
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:10}
\begin{aligned}
x_{ijk} \in \{0, 1\} \forall i,j,k;
\end{aligned}
\end{equation}

\begin{equation}
\label{eq:11}
\begin{aligned}
y_i \in \{0,1\} \forall i;
\end{aligned}
\end{equation}

@@

%-----------------------------------------------------------------------------%
\subsection{Evolution Algorithms}
\label{ssec:evolution-algorithms}
%-----------------------------------------------------------------------------%
\textit{Evolution Algorithms (EAs)} merupakan salah satu \textit{family} dari algoritma optimasi yang terinspirasi dari alam. Pada algoritma evolusi, terdapat proses seleksi dan reproduksi untuk memurnikasi populasi dari kandidat solusi \cite{eugster_many_2003}. Secara umum \textit{lifecycle} dari \textit{Evolution Algorithms} diawali dengan  konfigurasi populasi awal secara random. Kemudian, setiap iterasi dimulai dengan evaluasi atas fungsi obyektif atas individu pada populasi. Berdasarkan hasil evaluasi, kemudian \textit{fitness value} disematkan pada setiap kandidat solusi pada populasi. Kandidat solusi tersebut kemudian mengalami reproduksi dan menghasilkan generasi berikutnya. Proses kemudian berulang dengan menggunakan individu generasi yang baru.

Coevolutionary (CoE) algorithms (CoEA), merupakan pengembangan dari EAs. In standard EAs, evolution is usually viewed as if the population attempts to adapt in a fixed physical environment. In contrast, coevolutionary (CoE) algorithms (CoEA) realize that in natural evolution the physical environment is influenced by other independently-acting biological populations \cite{engelbrecht_coevolution_2007}. CoEA dapat dibagi menjadi dua, tergantung interaksi dari masing-masing spesies: \textit{competitive} dan \textit{cooperative}. Pada \textit{competitive coevolution} setiap individu berkompetisi dengan satu kelompoknya. Sementara pada \textit{cooperative coevolution}, interaksi antar spesies bersifat saling menguntungkan, atau paling tidak, tidak saling membahayakan \cite{engelbrecht_coevolution_2007}.

%-----------------------------------------------------------------------------%
\subsection{Publish/Subscribe Paradigm}
\label{ssec:pub-sub}
%-----------------------------------------------------------------------------%
\textit{Publish/subscribe interaction} merupakan salah satu metode komunikasi antara \textit{client} dan \textit{server}. Paradigma interaksi pada \textit{publish/subscribe} adalah adanya \textit{subscriber} yang memiliki ketertarikan pada suatu \textit{event} atau \textit{pattern of event}, agar dapat dikirimkan notifikasi tentang sebuah \textit{event} oleh \textit{publisher} yang sesuai dengan \textit{interest}-nya \cite{eugster_many_2003}.

Model dasar dari sistem \textit{publish/subscribe}, seperti Gambar \ref{fig:pub-sub-general}, bergantung pada \textit{event notification service} yang menyediakan penyimpanan dan \textit{management of subscription}. \textit{Event service} tersebut berperan sebagai mediator antara \textit{publisher} yang berperan sebagai \textit{producer} dari \textit{event} dan \textit{subscriber} yang berperan sebagai \textit{consumer} dari \textit{event}.

\begin{figure}[h]
	\centering
	\includegraphics[width=8cm]{Resources/Images/pub-sub-general}
	\caption{Arsitektur dasar pada Pub/Sub \cite{eugster_many_2003}}
	\label{fig:pub-sub-general}
\end{figure}

Salah satu keuntungan dari mekanisme publish/subscribe adalah bersifat loose coupling \cite{eugster_many_2003} antara publisher dengan subscriber. Pemisahan informasi (\textit{decoupling}) yang terjadi antara \textit{subscriber} dan \textit{publisher} dapat dipisahkan dalam 3 (tiga) dimensi, yaitu:

\begin{enumerate}
\item \textit{Space decoupling.} \\
Interaksi antara \textit{publisher} dan \textit{subscriber} tidak perlu saling mengetahui satu dengan yang lain. \textit{Publisher} mengirimkan \textit{event}, dan \textit{subscriber} menerima \textit{event} secara tidak langsung melalui \textit{event service}. \textit{Publisher} biasanya tidak memegang referensi terhadap \textit{subscriber}. Begitu juga sebaliknya, \textit{subscriber} biasanya tidak memegang referensi terhadap \textit{publisher}.
\item \textit{Time decoupling.} \\
Pihak-pihak yang berinteraksi, tidak harus berinteraksi pada waktu yang bersamaan. \textit{Publisher} dapat mengirimkan \textit{event} pada saat \textit{subscriber} dalam kondisi terputus. Begitu juga sebaliknya, \textit{subscriber} tetap dapat menerima \textit{event}, meskipun \textit{original publisher} dalam kondisi terputus.
\item \textit{Synchronizing decoupling.} \\
\textit{Publisher} tidak diblok ketika memproduksi \textit{event}, dan juga \textit{subscriber} tetap dapat memperoleh informasi meskipun sedang mengerjakan tugas yang lain. \textit{Publisher} dan \textit{subscriber} tidak berada dalam \textit{main flow}, sehingga tidak berinteraksi secara \textit{synchronous}.
\end{enumerate}

%-----------------------------------------------------------------------------%
\section{Proposed Solution}
\label{sec:proposed-solution}
%-----------------------------------------------------------------------------%
Kelemahan MDVRP ketika digunakan dalam rekomendasi lokasi pencacahan adalah ketiadaan data service time, padahal data itu sangat penting dalam kalkulasi rekomendasi. Untuk itu, MDVRP perlu diintegrasikan dengan mekanisme real-time. Terdapat beberapa mekanisme yang dapat diadopsi pada real-time system, diantaranya: Web service, RPC, message passing, dan publish/subscribe \cite{eugster_many_2003}.

Mekanisme publish/subscribe digunakan dalam penelitian ini karena memiliki sifat \textit{loose coupling}. Selain itu mekanisme publish/subscribe juga sesuai sesuai digunakan pada sistem yang bersifat information driven \cite{muhl_large-scale_2002}. Karena karakteristiknya ini, maka publish/subscribe dapat bekerja secara asynchronous, dimana \textit{request} dan \textit{reply} tidak harus diproses secara berurutan. Pada saat \textit{subscriber} melakukan subscribe, publisher tidak harus dalam kondisi online, begitu juga sebaliknya, pada saat \textit{publisher} melakukan publish, \textit{subscriber} tidak harus dalam kondisi online.

\begin{figure}[h]
	\centering
	\includegraphics[width=8cm]{Resources/Images/system-overview}
	\caption{Garis Besar Sistem}
	\label{fig:system-overview}
\end{figure}

Gambar \ref{fig:system-overview} menunjukkan garis besar sistem, dimana sistem tersusun atas 3 (tiga) komponen utama, yaitu: \textit{publisher}, \textit{subscriber}, dan \textit{message broker}. Komunikasi antara \textit{publisher} dan \textit{subscriber} didasari atas satu kesamaan, yaitu \textbf{event} atau \textbf{topic}. Topik yang digunakan sebagai dasar komunikasi antara publisher dan subscriber dalam penelitian ini adalah \textit{current location} dari masing-masing subscriber.

%-----------------------------------------------------------------------------%
\subsection{Recommendation Publisher}
\label{ssec:recommendation-publisher}
%-----------------------------------------------------------------------------%
VRP Solver diimplementasikan pada sisi publisher, yang akan melakukan pencarian solusi/rute dan kemudian dipublish kepada \textit{subscriber} melalui \textit{message broker}. Sebagaimana sebelumnya disebutkan, bahwasannya \textit{current location} dari masing-masing \textit{subscriber} digunakan sebagai topik. Sebuah thread disiapkan, yang secara periodik akan melakukan pengecekan apakah terdapat topik yang baru pada message broker.

Setiap kali terdapat topik yang baru, sebuah thread akan disiapkan, dan topik tersebut digunakan sebagai ID dari thread yang bersangkutan. Pada thread tersebut terdapat procedure VRP solver yang akan melakukan pencarian solusi/rute. Untuk itu, sebuah threadpool disiapkan untuk menampung thread dari masing-masing topik tersebut secara berurutan berdasarkan kedatangan. Karena topik digunakan sebagai thread ID, maka ukuran dari threadpool akan sama dengan jumlah seluruh lokasi pencacahan.

Setiap thread dalam threadpool kemudian dijalankan secara berurutan, satu session untuk masing-masing thread. Setiap thread yang dijalankan, yang menggunakan topik sebagai ID, akan melakukan pencarian solusi dengan melibatkan seluruh $M$ vehicles dan $(unassigned) N$ locations. Pencarian solusi dilakukan pada procedure VRP solver pada section \ref{ssec:vrp-solver}.

\begin{algorithm}[h]
	\caption{TopicWatcher}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}}
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\REQUIRE $C$
		\ENSURE  $T$
		\FOR {$i = 1$ to $len(C)$}
			\FOR {$j = 1$ to $N$}
				\IF {($C_i == E_j$)}
					\STATE $T_j = Thread(C_i, V_1...V_M, (unassigned) E_1...E_N)$
				\ENDIF
			\ENDFOR
		\ENDFOR
		\RETURN $T$
	\end{algorithmic}
\end{algorithm}

Proses dalam procedure VRP solver pada masing-masing thread akan menghasilkan sejumlah minimal $1$ rute, dan maksimal $M$ rute. Setiap rute $R_i$ yang dihasilkan kemudian akan dipublish dengan topik $C_i$. Setelah rute dipublish, \textit{message broker} akan menginformasikan kepada publisher, jumlah \textit{subscriber} yang menerima pesan. Adapun identitas \textit{subscriber} yang menerima pesan tetap tidak diketahui oleh \textit{publisher}.

Setiap rute $R_i$ yang memiliki jumlah penerima pesan lebih dari satu, maka \textit{thread} yang mempunyai ID $C_{R_i}$ akan dicancel. Pembatalan \textit{thread} ini dimaksudkan agar solusi/rute yang telah diterima oleh \textit{subscriber} tidak dikalkulasi kembali. Terdapat sebuah kondisi, dimana pada saat thread menjalankan VRPSolver dengan ID $C_i$, ternyata rute yang dihasilkan tidak mengandung solusi untuk topik $C_i$. Pada kondisi tersebut, maka topik $C_i$ tersebut akan diantrikan kembali dengan hanya menyertakan dirinya sendiri sebagai pencacah, sehingga diperoleh kepastian topik $C_i$ akan memperoleh solusi/rute.

@@

Salah satu keuntungan mekanisme publish/subscribe, yang membuat dia fleksible, adalah karakteristiknya yang loose coupling. Dalam kasus ini, keuntungan dari mekanisme ini sekaligus menjadi kelemahan. Ketidaktahuan \textit{publisher} akan identitas \textit{subscriber} menyebabkan \textit{current location} dari keseluruhan \textit{subscriber} tidak dapat diketahui. Untuk itu diperlukan mekanisme yang dapat mengatasi permasalahan pertukaran identitas, yaitu dengan menggunakan \textit{shared memory}. Pada \textit{shared memory} tersimpan data \textit{current location} dari seluruh \textit{subscriber}, yang mana \textit{publisher} dapat mengetahui \textit{current location} dari seluruh \textit{subscriber} dari data yang tersimpan pada \textit{shared memory}.

Proses tersebut diatas terus berulang, sampai dengan seluruh lokasi pencacahan semuanya ter-assign. Gambar \ref{fig:publisher-algorithm} memberikan ilustrasi algoritma yang digunakan pada \textit{recommendation publisher}.

@@

Masing-masing rute kemudian dihitung \textit{total cost}-nya, yang merupakan penjumlahan dari seluruh \textit{service time} pada masing-masing lokasi dan seluruh waktu tempuh dari masing-masing perpindahan. Metric yang akan dijadikan pembanding adalah standar deviasi dari seluruh rute pada masing-masing program dan program pembanding. Program yang lebih baik akan menghasilkan standar deviasi yang lebih kecil. Standar deviasi dipilih sebagai \textit{metric} karena merepresentasikan kondisi sebenarnya dilapangan, dimana jika variasi waktu dari seluruh pencacah lebih kecil, maka penyelesaian pencacahan akan lebih merata.

>>>>

%
%
\usepackage[bookmarks=false]{hyperref}
\newcommand{\algorithmautorefname}{Algorithm}

%
%
\usepackage{minted}
\usepackage{tcolorbox}
\usepackage{etoolbox}
\BeforeBeginEnvironment{minted}{\begin{tcolorbox}}%
\AfterEndEnvironment{minted}{\end{tcolorbox}}%
\AtBeginEnvironment{minted}{\fontsize{8}{8}}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{An Architecture for Real-Time Location Recommendation for Field Data Collection}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Field data collection is one of the main activities performed by the national statistical agencies in every country. Data collection activities have a similiar workflow with Multi Depot Vehicle Routing Problem (MDVRP). The use of MDVRP to generate pre-calculated routes resulted in a total route costs with high standard deviation. Real-time mechanism by utilizing the publish/subscribe paradigm combined with Cooperative Coevolution Algorithms (CoEAs) is proposed to reduce the inequality (large variation) of the completion time. The test results show that the routes produced by publish/subscribe paradigm combined with CoEAs are more prevalent in total route times among enumerators compared with pre-calculated routes produced by only CoEAs.
\end{abstract}

@@

%-----------------------------------------------------------------------------%
\section{Introduction}
\label{sec:introduction}
%-----------------------------------------------------------------------------%
Data collection is one of the main duties of the statistical agency in every country, including BPS-Statitics Indonesia. There are two data collection methods that are generally used: primary data collection and secondary data collection. In the primary data collection, the enumerators directly meet and interview the respondents, while in the secondary data collection, the statistical institution only compiles the data collected by other agencies. The primary data collection method requires the enumerators to be allocated to several census/survey areas that have been specified before. The allocation is usually made by assigning an equal number of locations on each enumerator.

The workflow of the data collection performs as follows: 
\begin{enumerate}
	\item The enumeration will start on a particular point. It could be the statistical office or the enumerator's house. 
	\item The enumerator moves to the first census/survey location and visits all respondents in that location. 
	\item After finishing the first census/survey area, the enumerator moves to the second census/survey location and continue doing so until all locations have been visited successfully.
\end{enumerate}

This workflow is significantly similar to the Vehicle Routing Problem (VRP), or more specifically, the Multi Depot Vehicle Routing Problem (MDVRP).

However, although the data collection shares similar workflow with MDVRP, the rough implementation of the MDVRP algorithm is not feasible to solve the allocation problems. This is because one important variables needed in the algorithm, the service time, is not available until after the enumerator finish visiting the census/survey location. Therefore, this research aims to propose an integration of MDVRP with pub/sub mechanism to create a location recommendation system for field data collection that can be used in real time.

%-----------------------------------------------------------------------------%
\section{Literature Review}
\label{sec:literature-review}
%-----------------------------------------------------------------------------%
%-----------------------------------------------------------------------------%
\subsection{Multi-Depot Vehicle Routing Problem}
\label{ssec:mdvrp}
%-----------------------------------------------------------------------------%
Multi-Depot Vehicle Routing Problem (MDVRP) is one of classic  Vehicle Routing Problem (VRP) variance, where there are more than one depot that can be used  \cite{montoya-torres_literature_2015}. \autoref{fig:mdvrp-illustration} shows an example of the solution for VRP problem that uses two depot and two vehicle routes which associated with each depot. Basically, a solution to this problem is a set of vehicle routes such that: (i) each vehicle route starts and ends at the same depot, (ii) each customer is served exactly once by one vehicle, (iii) the total demand on each route does not exceed vehicle capacity (iv) the maximum route time is satisfied and (v) the total cost is minimized.

\begin{figure}[h]
	\centering
	\includegraphics[width=6cm]{Resources/Images/mdvrp-illustration}
	\caption{An illustration for \textit{Multi Depot} VRP}
	\label{fig:mdvrp-illustration}
\end{figure}

According to Renaud et al \cite{renaud_tabu_1996}, the MDVRP can be formally described as follows. Let $G = (V, A)$ be a complete graph, where $V$ is the set of nodes and $A$ is the set of arcs. The nodes are partitioned into two subsets: the customers to be served, $V_C$ = \{1,..., $N$\}, and the multiple depots $V_D$ = \{$N$+1,..., $N+M$\}, with $V_C \cup V_D$ = $V$ and $V_C \cap V_D$ = $\oslash$. There is a non-negative cost $c_{ij}$ associated with each arc $(i, j) \in A$. The demand of each customer is $d_i$ (there is no demand at the depot nodes). There is also a fleet of $K$ identical vehicles, each with capacity $Q$. The service time at each customer $i$ is $t_i$ while the maximum route duration time is set to $T$. $A$ conversion factor $w_{ij}$ might be needed to transform the cost $c_{ij}$ into time units. In the classical MDVRP, however, the cost is the same as the time and distance units, so $w_{ij}$ = $1$.

In the mathematical formulation \cite{kulkarni_integer_1985}, binary variables $x_{ijk}$ are equal to $1$ when vehicle $k$ visits node $j$ immediately after node $i$. Auxiliary variables $y_i$ are also used in the subtour elimination constraints.

Minimize
\begin{flalign}
\label{eq:1}
\sum_{i=1}^{N+M}\sum_{j=1}^{N+M}\sum_{k=1}^{K}c_{ij}x_{ijk};
\end{flalign}

Subject to:
\begin{flalign}
\label{eq:2}
\sum_{i=1}^{N+M}\sum_{k=1}^{K}x_{ijk} = 1  (j=1,..., N);
\end{flalign}

\begin{flalign}
\label{eq:3}
\sum_{j=1}^{N+M}\sum_{k=1}^{K}x_{ijk} = 1  (j=1,..., N);
\end{flalign}

\begin{flalign}
\label{eq:4}
&\sum_{i=1}^{N+M}x_{ihk} - \sum_{j=1}^{N+M}x_{hjk} = 0 \\
\nonumber
&(k=1,...,K; h=1,...,N+M);
\end{flalign}

\begin{flalign}
\label{eq:5}
\sum_{i=1}^{N+M} \sum_{j=1}^{N+M} d_ix_{ijk} \leq Q (k=1,...,K);
\end{flalign}

\begin{flalign}
\label{eq:6}
\sum_{i=1}^{N+M} \sum_{j=1}^{N+M} (c_{ij}w_{ij} + t_i) x_{ijk} \leq T (k=1,...,K);
\end{flalign}

\begin{flalign}
\label{eq:7}
\sum_{i=N+1}^{N+M} \sum_{j=1}^{N} x_{ijk} \leq 1 (k=1,...,K);
\end{flalign}

\begin{flalign}
\label{eq:8}
\sum_{j=N+1}^{N+M} \sum_{i=1}^{N} x_{ijk} \leq 1 (k=1,...,K);
\end{flalign}

\begin{flalign}
\label{eq:9}
&y_i - y_j + (M + N)x_{ijk} \leq N + M - 1; \\
\nonumber
&for 1 \leq i \neq j \leq N and 1 \leq k \leq K;
\end{flalign}

\begin{flalign}
\label{eq:10}
x_{ijk} \in \{0, 1\} \forall i,j,k;
\end{flalign}

\begin{flalign}
\label{eq:11}
y_i \in \{0,1\} \forall i;
\end{flalign}

@@

%-----------------------------------------------------------------------------%
\subsection{Evolution Algorithms}
\label{ssec:evolution-algorithms}
%-----------------------------------------------------------------------------%
Evolution Algorithms (EAs) is one of the optimization algorithm families that are inspired by nature. In the evolution algorithm, there are selection and reproduction processes to purify the population from the solution candidates. According to Engelbreacht \cite{engelbrecht_coevolution_2007}, in general, the EAs lifecycle begins with randomly configuring the initial population. Afterwards, for each iteration, every individual in the population is evaluated using an objective function. Then, each solution candidate will be labeled with a fitness value based on their evaluation result. These solution candidates will then commit a reproduction and create the next generation. The whole process will be repeated using the new-generation individuals. These iterations will stop once it hits a certain time limit, which is 60 seconds.

Coevolutionary algorithms (CoEAs) is a development of EAs. In standard EAs, evolution is usually viewed as the population attempts to adapt in a fixed physical environment. In contrast, coevolutionary algorithms (CoEAs) realize that in natural evolution the physical environment is influenced by other independently-acting biological populations \cite{engelbrecht_coevolution_2007}. Based on the interaction of each species, CoEAs can be distinguished into two categories: competitive and cooperative \cite{engelbrecht_coevolution_2007}. In the competitive coevolution, each individual competees with other individuals in the same group. Meanwhile, in cooperative coevolution, the interaction between species is mutualistic, or at least, not harming each other.

%-----------------------------------------------------------------------------%
\subsection{Publish/Subscribe Paradigm}
\label{ssec:pub-sub}
%-----------------------------------------------------------------------------%
Publish/subscribe interaction is one of the communication methods between client and server. The interaction paradigm in publish/subscribe is the presence of subscribers that have an interest in a certain event or the pattern of an event. These subscribers will be notified by the publisher about the event they are interested in \cite{eugster_many_2003}.

The basic model of the publish/subscribe system, as can be seen in \autoref{fig:pub-sub-general}, depends on the event notification service that provides storage and management of subscription. This event service acts as a mediator between the publisher, who acts as the events producer, and the subscriber who acts as the event consumer.

\begin{figure}[h]
	\centering
	\includegraphics[width=8cm]{Resources/Images/pub-sub-general}
	\caption{Basic architecture in Pub/Sub \cite{eugster_many_2003}}
	\label{fig:pub-sub-general}
\end{figure}

One of the benefits of publish/subscribe mechanism is the loose coupling characteristic \cite{eugster_many_2003} between the publisher and the subscriber. The separation of information (decoupling) between publisher and subscriber can be divided into 3 (three) dimension, which are:

\begin{enumerate}
	\item Space decoupling. \\
	In doing the interaction, both publisher and subscriber do not need to know each other. The publisher sends an event and subscriber indirectly receive the event through the event service. The publisher does not hold any reference regarding the subscriber and vice versa. 
	\item Time decoupling. \\
	The parties who are involved in the interactions do not have to interact in the same time. The publisher can send the event even though he is in a disconnected state. Similarly, the subscriber can still receive the event, albeit disconnected. 
	\item Synchronizing decoupling. \\
	Publisher will not be blocked when producing the event. The subscriber can still obtain information despite, at the same time, working on other tasks. Publisher and subscriber are not in the main flow, hence are not interacting synchronously. 
\end{enumerate}

%-----------------------------------------------------------------------------%
\section{Proposed Solution}
\label{sec:proposed-solution}
%-----------------------------------------------------------------------------%
The drawback of using MDVRP for location recommendation in field data collection is the absence of time service data, which is very important in the recommendation calculation. Thus, MDVRP needs to be integrated with the real-time mechanism. There are several mechanisms that can be adopted in a real-time system such as Web Service, Remote Procedure Call (RPC), message passing, and publish/service \cite{eugster_many_2003}.

The publish/subscribe mechanism is used in this research because it has the \textit{loose coupling} characteristic. Furthermore, publish/subscribe mechanism is also suitable for an information-driven system \cite{muhl_large-scale_2002}. Because of this characteristic, the publish/subscribe mechanism can work asynchronously, where request and reply do not have to be processed in sequence. The publisher and subscriber can conduct an offline publish and offline subscribe, respectively.

\begin{figure}[h]
	\centering
	\includegraphics[width=8cm]{Resources/Images/system-overview}
	\caption{The system outline}
	\label{fig:system-overview}
\end{figure}

\autoref{fig:system-overview} shows the system outline that consists of 3 (three) main components: the publisher, the subscriber, and the message broker. The communication between publisher and subscriber is built upon a similarity in either event or topic. In this research, the topic that is used as the basis for the communication between publisher and subscriber is the \textit{current location} of each subscriber.

%-----------------------------------------------------------------------------%
\subsection{Recommendation Publisher}
\label{ssec:recommendation-publisher}
%-----------------------------------------------------------------------------%
VRP solver is implemented on the publisher's side, who will search for solution/route and then publish it to the subscriber via the message broker. The subscriber's current location is used as a topic. Afterward, a thread  (\autoref{alg:topic-watcher}) is prepared that will periodically check if there is any new topic at message broker.

Each time a new topic emerge, a new thread will be created and that new topic will be used as an ID for that thread. Each thread contains VRP solver procedure (\autoref{alg:vrp-worker}) that will do the solution/route finding. Accordingly, a threadpool is provided to contain all threads from each topic in order of arrival time. Since topic is used as a thread ID, the threadpool size is equal to the number of census/survey location.

Each thread in thread pool will be run in order, one session for each thread. Every running thread, with a topic as an ID, will find the solution by embroiling all $M$ vehicles and $(unassigned) N$ locations. The solution finding is done using VRP solver procedure which is explained in more detail in \autoref{ssec:vrp-solver}.

\begin{algorithm}[h]
	\caption{TopicWatcher}
	\label{alg:topic-watcher}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}}
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\REQUIRE $C$
		\ENSURE  $T$
		\FOR {$i = 1$ to $len(C)$}
			\FOR {$j = 1$ to $N$}
				\IF {($C_i == E_j$)}
					\STATE $T_j = Thread(C_i, V_1...V_M, (unassigned) E_1...E_N)$
				\ENDIF
			\ENDFOR
		\ENDFOR
		\RETURN $T$
	\end{algorithmic}
\end{algorithm}

The process in VRP solver procedure in each thread will produce at least $1$ route and at most $M$ routes. Every route $R_i$ that is produced will be published with topic $C_i$. After being published, the message broker will inform the publisher about the number of subscribers that receive the message. The identity of the subscribers that receive the message will remain unknown to the publisher.

If a route $R_i$ has more than one message receiver, the thread with ID $C_{R_i}$ will be cancelled. This annulment is aimed to prevent a solution/route that has been received by subscribers from being recomputed. There is a condition when the running VRP solver with ID $C_i$ does not produce the solution for topic $C_i$. In this kind of situation, the topic $C_i$ will be requeued with only including the subscriber as the enumerator. By doing this, there is a guarantee that the topic $C_i$  will obtain a solution/route.

@@

As mentioned before, the advantage of using publish/subscribe mechanism is its loose coupling characteristic that makes it flexible. But this benefit could also be deemed as a shortcoming. The publisher's unawareness towards subscribers' identities has caused the current location of all subscribers cannot be identified. Hence, we need a mechanism that can support information exchanging between publisher and subscribers. The idea is to use a shared memory contains the current location data of all subscribers. By accessing this shared memory, the publisher will be able to determine the subscribers' current locations.

The process above will keep repeating until all enumeration locations have been assigned to the subscribers (enumerators). \autoref{fig:publisher-algorithm} illustrates the workflow of the algorithm used in recommendation publisher.

@@

The total cost for each route is calculated by adding up the service time and the travel time from all locations in that route. The standard deviation of all routes in each program (the proposed program and the benchmark program) is used as a metric for comparing the outputs. A better program will result in a smaller standard deviation. Standard deviation is chosen as the metric because it represents the real field condition. If the time variance between each enumerator is smaller, the enumeration completion will be equal and fair.
